An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-23 draft.tex
A few changes

% There is a big gap here where you go from generic considerations
% about cloud computing to the presentation of this method. In the
% middle, there should be motivation for it, what are you trying to
% address precisely, and how you intend to address it. "Approaching
% multi-population algorithms in cloud computing" is not nearly
% enough. There are several points here, and every one of them needs
% to be addressed and justified
% * Using containers as opposed to a serverless architecture or a
% virtual-machine based one.
% * Combining several metaheuristics instead of using a single one.
% * Reducing the amount of parameters.
% There needs to be a motivation for every one, relate them to the
% objectives of the paper (which need to be clarified) and eventually
% prove the contribution of every one of them to the final result - JJ

@@

% Based on a reactive architecture \cite{guervos2018introducing} , each
% subpopulation is represented as an event message that is pushed asynchronously
% into a message queue. Incoming messages trigger stateless functions that receive
% the subpopulation and proceed to run an algorithm, using the parameters and
% population included in the message. After the specified number of iterations,
% each stateless function returns the evolved subpopulation by again pushing a
% message to another queue, used for receiving the resulting subpopulations.
% Subpopulations are received from the queue by a controller that is responsible
% for mixing the individuals from different subpopulations and producing new
% subpopulations. These new subpopulations are pushed again by the controller into
% the message queue, creating a loop. The cycle stops when the controller receives
% a subpopulation containing a candidate solution that satisfies a particular
% condition, or a maximum number of messages were received.

@@

The most elemental data structure used throughout the system is the subpopulation; and at a higher
level, we can view the system as a continuous stream of subpopulations
flowing
% We talk about "multi-population" elsewhere, including the
% abstract. We should unify terminology - JJ
% "multi-population" is the whole thing, a subpopulation is a member of that.
% "A multi-population has many subpopulations or should we call them just populations?"
% I will add a comment about this in the Multi-population section -
% Mario
% use the usual "multi-deme" terminology. A population has different
% sub-populations nad uses a multi-deme implementation - JJ
from one process to the other. By "continuous stream," we mean that ideally,
everything must happen asynchronously without components needing to wait for
others. In this kind of system, this streaming functionality is normally
implemented by using a message queue system. For instance, if one process needs
to communicate a subpopulation to another, it must pack the sub-population into
a message, and then send the message asynchronously % I don't know if
                                % asynchronously applies here. It's
                                % simply sent - JJ 
                                % Meaning it does not wait for a
                                % respone, it continues doing other
                                % stuff - Mario
% That's rather "fire and forget". Or simply say it's sent without
% waiting for a response - JJ
to the queuing system; this
means that after the process pushes the message, it continues its execution
without waiting for a result. On the other end, the recipient subscribes to the
message queue by defining an event handler method that is going to be triggered when a
message is pushed to it. The two components at each end of the queue are
Producers that push messages to the queue and Consumers that pull them. This
pattern is an essential component of a reactive architecture and makes
it 
highly scalable; one of the reasons for this is the use of functions
that have no secondary effects or access global storage, that is,
stateless functions. These functions do not need to read, keep, or modify data outside of the method. If
Consumers are implemented using stateless functions, and thus, there is no difference
between having one or many copies of the same function pulling work from the queue all at the
same time, because no side effects, including unwanted resource
locking problems,  are going to result from this concurrency.
% This is important and we should probably mention that in the
% introduction - JJ 
Based on these general principles we define next the proposed model.

@@

The algorithm starts with the Setup process pulling a configuration message from
the experiements queue, this is not shown in the diagram. The configuration messsage 
includes all the parameters needed to execute the algorithm, among other things,
the number of subpopulations, the number of individuals, and the number of iterations
of the algorithm. We give more details about the configuration data structure
in Section []. % Missing
Once the configuration is read, we can follow the path of 
messagess as follows:

@@

We have mentioned several advantages of a cloud-native application for running
multi-population based algorithms, but there are in our experience several
practical issues of this approach. In general, cloud platforms have free-tiers
for their services, and many offer academic discounts to students. But once we
pass to the pay-as-you-go tier, computing resources need to be
managed, and, where possible, its cost minimized.  First,
in many cases,  a credit-card is needed to guarantee the payment of the
resources consumed and someone responsible for assigning quotas and, more
importantly, special keys for using the cloud's API.  If these keys are
compromised or shared by their holders,  additional costs or administrative
procedures are in order. For those cases when a prototype platform is needed,
for instance, students developing a new algorithm variant or experimenting with
a published work, a container-based platform can be more practical. Container
technology enables researchers the local deployment of all the infrastructure
needed to run the experiments. Later, if needed, the same deployment could be
deployed to the cloud.
% I don't see this line of argument. You can do exactly the same with
% virtual machines instead of containers. Also, you can test a virtual
% infrastructure locally and then only when it's needed deploy it to
% the cloud. If you want to give good reasons for using containers
% instead of virtual machines, there are quite a lot non-economic ones
% - JJ

@@

\subsection{Redis Container} 
\label{redis}

% Do we need a description of every container? Wouldn't a general
% description of the containers used be enough? - JJ

@@

The controller is an essential component of the architecture (Figure \ref{fig:kafkEO})
because it is responsible for maintaining the evolutionary loop. It takes newly
evolved populations from the Output Queue, mixes them,   and produces new
messages from the result. At the same time, it must keep track of two conditions
for ending the loop: the number of messages it has received reaches a maximum
value, or the error of the best solution goes below a specific threshold.  Finally, it
must filter out remnant messages from other experients. Messages from other
experiments can remain in the queues because of the asynchronous nature of the
system.  We chose to implement the controller in Python but using an API
specialized in asynchronous event-driven programming over data streams.  The
open-source library is called Reactive Extensions for Python (RxPY)
\url{https://github.com/ReactiveX/RxPY} based on the Observer pattern
\cite{gamma1995design} and functional programming.

>>>>

@@

@@

The most elemental data structure used throughout the system is the subpopulation; and at a higher
level, we can view the system as a continuous stream of subpopulations
flowing
% We talk about "multi-population" elsewhere, including the
% abstract. We should unify terminology - JJ
% "multi-population" is the whole thing, a subpopulation is a member of that.
% "A multi-population has many subpopulations or should we call them just populations?"
% I will add a comment about this in the Multi-population section -
% Mario
% use the usual "multi-deme" terminology. A population has different
% sub-populations nad uses a multi-deme implementation - JJ
% Let's talk about it, I am using the Multi-population terminology 
% maybe multi-dem is more common in the EA community but not 
% much in the Swarm Intelligence side?  
% Any how, we decide and explain the difference so readers of both 
% camps are Ok with it. 
% I prefer the use of population and something like the global population - Mario

from one process to the other. By "continuous stream," we mean that ideally,
everything must happen asynchronously without components needing to wait for
others. In this kind of system, this streaming functionality is normally
implemented by using a message queue system. For instance, if one process needs
to communicate a subpopulation to another, it must pack the sub-population into
a message, and then send the message 
% I don't know if
                                % asynchronously applies here. It's
                                % simply sent - JJ 
                                % Meaning it does not wait for a
                                % respone, it continues doing other
                                % stuff - Mario
% That's rather "fire and forget". Or simply say it's sent without
% waiting for a response - JJ
% corrected - Mario
to the queuing system without waiting for a response; this
means that after the process pushes the message, it continues its execution
without waiting for a result. On the other end, the recipient subscribes to the
message queue by defining an event handler method that is going to be triggered when a
message is pushed to it. The two components at each end of the queue are
Producers that push messages to the queue and Consumers that pull them. This
pattern is an essential component of a reactive architecture and makes
it 
highly scalable; one of the reasons for this is the use of functions
that have no secondary effects or access global storage, that is,
stateless functions. These functions do not need to read, keep, or modify data outside of the method. If
Consumers are implemented using stateless functions, and thus, there is no difference
between having one or many copies of the same function pulling work from the queue all at the
same time, because no side effects, including unwanted resource
locking problems,  are going to result from this concurrency.
% This is important and we should probably mention that in the
% introduction - JJ 
Based on these general principles we define next the proposed model.

@@

The algorithm starts with the Setup process pulling a configuration message from
the experiements queue, this is not shown in the diagram. The configuration messsage 
includes all the parameters needed to execute the algorithm, among other things,
the number of subpopulations, the number of individuals, and the number of iterations
of the algorithm. We give more details about the configuration data structure
in Section \ref{experiment_flow}. % Missing  % added 
Once the configuration is read, we can follow the path of 
messagess as follows:

@@

We have mentioned several advantages of a cloud-native application for running
multi-population based algorithms, but there are in our experience several
practical issues of this approach. In general, cloud platforms have free-tiers
for their services, and many offer academic discounts to students. But once we
pass to the pay-as-you-go tier, computing resources need to be
managed, and, where possible, its cost minimized.  First,
in many cases,  a credit-card is needed to guarantee the payment of the
resources consumed and someone responsible for assigning quotas and, more
importantly, special keys for using the cloud's API.  If these keys are
compromised or shared by their holders,  additional costs or administrative
procedures are in order. For those cases when a prototype platform is needed,
for instance, students developing a new algorithm variant or experimenting with
a published work, a container-based platform can be more practical. Container
technology enables researchers the local deployment of all the infrastructure
needed to run the experiments. Later, if needed, the same deployment could be
deployed to the cloud.
% I don't see this line of argument. You can do exactly the same with
% virtual machines instead of containers. Also, you can test a virtual
% infrastructure locally and then only when it's needed deploy it to
% the cloud. If you want to give good reasons for using containers
% instead of virtual machines, there are quite a lot non-economic ones
% - JJ
% OK, I will add an issue - Mario

@@

% Do we need a description of every container? Wouldn't a general
% description of the containers used be enough? - JJ 
% Ok, I think only workers and the controller need to be described in detail - Mario

@@

The controller is an essential component of the architecture (Figure \ref{fig:kafkEO})
because it is responsible for maintaining the evolutionary loop. It takes newly
evolved populations from the Output Queue, mixes them,   and produces new
messages from the result. At the same time, it must keep track of two conditions
for ending the loop: the number of messages it has received reaches a maximum
value, or the error of the best solution goes below a specific threshold.  Finally, it
must filter out remnant messages from other experients. Messages from other
experiments can remain in the queues because of the asynchronous nature of the
system.  We chose to implement the controller in Python but using an API
specialized in asynchronous event-driven programming over data streams.  The
open-source library is called Reactive Extensions for Python (RxPY)
\url{https://github.com/ReactiveX/RxPY}an it is based on the Observer pattern
\cite{gamma1995design} and functional programming.
